Discovering Robustness Threats by Mining the Exception Stack Traces of over 600 Android Projects

### 2014-09-03 android-stacks.tex
addressing comments. introduction2

@@

@@

In Java, when an application fails due to an uncaught exception, 
it automatically terminates, while the system prints a stack trace to the console, 
or log file ~\cite{gosling2000java}.  A typical Java stack trace consists of  the fully qualified name 
the thrown exception and the ordered list of methods that were active on the call stack before 
such exception has occurred.

@@

Most of such studies, however, focused the analysis on one or two system at a time.
This paper reports an exploratory study which targets a family of applications 
(sharing the common Android framework) aiming at extracting 
the common characteristics of exception stack traces reported to such systems to provide
insights on similar fragilities on the exception related code of the applications and underlaying framework.

@@

To do so we perform a post mortem analysis of the exception stack traces embedded on all issues
created on Android projects available in GitHub and Google Code. Using a custom tool called ExceptionMiner,
 which we develop specifically for this study, we mine stack traces from issues across all Android projects on
GitHub and Google Code. Overall, we analyzed all issues extracted from on 482 Android projects defined in
GitHub and 157 projects defined in Google Code. Overall it corresponds to 159,048 issues (31,592 issues from GitHub projects 
and 127,456 issues from Google Code projects) from which  6,005 stack traces were extracted and used as
 a source of information for our analysis. The stack trace analysis was augmented by additional information
extracted from JDK, Android Plaform and the applications using byte code and source code analysis.
 Manual inspection was used to leverage the understanding of exception stack traces and support further discussions and insights.

@@

Such fragilities represent obstacles to the development of dependable java-based applications as they make difficult and 
sometimes almost impossible to protect the code against the Uncaught Exceptions that lead to system crashes.
Although the study focused on Android open-source applications, these findings are not restricted to this context, since 
the above mentioned fragilities may happen in any Java-based system. 
For instance: the unexpected wrappings were performed by Java libraries and elements (e.g. reflection library and 
the static blocks); and the undocumented runtime exceptions were also found in Java libraries besides the Android platform.
The study results are relevant to: (i) Android and Java developers to familiarize common characteristics and fragilities of exception-related code 
which is the first step to help developers preventing and dealing with them; (ii) designers of languages and tools to consider ways of reducing 
such fragilities in Android and Java environments.

>>>>

%\title{Results of a Mining Study on Exception Stack Traces embedded on issues of over 600 Android Projects}

@@

%\title{Discovering Threats to Robustness through Exception Stack Trace Analysis: Results of a Mining Study of Android Projects}

%\title{Discovering Threats to Robustness through Exception Stack Trace Mining}

%\title{Discovering Threats to Robustness through Exception Stack Trace Mining of over 600 Android Projects}

%\title{Discovering Threats to Robustness through a Mining Study of Exception Stack Traces}

@@

In Java, when an application fails due to an uncaught exception, 
it automatically terminates, while the system prints a stack trace to the console, 
or log file ~\cite{gosling2000java}.  A typical Java stack trace consists of  the fully qualified name 
the thrown exception and the ordered list of methods that were active on the call stack before 
such exception has occurred ~\cite{gosling2000java,bloch2008effective}.

@@

%Most of such studies, however, focused the analysis on one or two system at a time.
%This paper reports an exploratory study which targets a family of applications 
%(sharing the common Android framework) aiming at extracting 
%the common characteristics of exception stack traces reported to such systems to provide
%insights on similar fragilities on the exception related code of the applications and underlaying framework.

@@

This paper reports a study which perform a post mortem analysis of the exception stack traces embedded on issues
reported on Android projects hosted in GitHub and Google Code. 
The goal of this study was to: investigate the whether the reported exception stack traces
 can reveal common fragilities on the exception-related code which may be favoring 
the introduction of failures such as uncaught exceptions - and as a consequence
 contributing to deacrease the robustness of apps.
To guide this investigation we compiled general guidelines on how to use
exceptions proposed by Gosling~\cite{gosling2000java},
Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective}.
Then, using a custom tool called ExceptionMiner,
 which we develop specifically for this study, we mine stack traces from the issues reported 
on 482 Android projects hosted in GitHub and 157 projects hosted in Google Code.
Overall 159,048 issues were analyzed and 6,005 stack traces were extracted from them.
The exception stack trace analysis was augmented with additional information
extracted from JDK, Android Plaform and the applications using byte code 
and source code analysis. Manual inspection was used to leverage the understanding 
of exception stack traces and support further discussions and insights.

@@

Such fragilities represent obstacles to the development of robust java-based applications as they make difficult and 
sometimes almost impossible to protect the code against the Uncaught Exceptions that lead to system crashes.
Although the study focused on Android open-source applications, these findings are not restricted to this context, since 
the above mentioned fragilities may happen in any Java-based system. 
For instance: the unexpected wrappings were performed by Java libraries and elements (e.g. reflection library and 
the static blocks); and the undocumented runtime exceptions were also found in Java libraries besides the Android platform.
The study results are relevant to: (i) Android and Java developers to familiarize common characteristics and fragilities of exception-related code 
which is the first step to help developers preventing and dealing with them; (ii) designers of languages and tools to consider ways of reducing 
such fragilities in Android and Java environments.
